---
title: "hw07"
author: "Victoria Michalowski"
date: "November 13, 2017"
output: github_document
---

For this assignment, I was having trouble downloading the gapminder data from the detailed instructions of the assignment, so I went to the gapminder website (http://www.gapminder.org/data/) looking for the original source, and found that they have many different kinds of worldwide data sets. I decided to choose one on CO2 emissions across different countries in the world.

The first step was to download the raw data, which I had already uploaded to my github repo.

```{r}
download.file("https://github.com/vmichalowski/STAT545-hw-Michalowski-Victoria/blob/master/hw07%20-%20automating%20pipelines/CO2_emissions-CDIAC.xlsx?raw=true", destfile="CO2_emissions-CDIAC.xlsx")
```

Next, I wanted to do some basic exploratory analysis.

Bring the data in as data frame.

```{r}
CO2 <- read.xlsx("CO2_emissions-CDIAC.xlsx", 
                sheetIndex = 1)
```

Save a couple descriptive plots to file with highly informative names.
Reorder the continents based on life expectancy. You decide the details.
Sort the actual data in a deliberate fashion. You decide the details, but this should at least implement your new continent ordering.
Write the Gapminder data to file(s), for immediate and future reuse.

Basic data cleaning:

Are there wonky factors? Consider bringing in as character, addressing their deficiencies, then converting to factor.
Are there variables you’re just willing to drop at this point? Do it!
Are there dates and times that need special handling? Do it!
Are there annoying observations that require very special handling or crap up your figures (e.g. Oceania)? Drop them!