---
title: "Gapminder Data Exploration"
author: "Victoria Michalowski"
date: "September 18, 2017"
output: github_document
---

#Exploring Gapminder Data

## This is the extension and polishing of R markdown documentation started in class on September 14, 2017.

## The first step is to load in the gapminder data.
```{r}
library(tidyverse)
library(gapminder)
summary(gapminder)
str(gapminder) #overview of the data frame, displays the structure of the object
gapminder #because of tidyverse, can print the data frame to the screen without cluttering the console
```
## Next, understand the data frames by inspecting the number of rows and columns.
```{r}
class(gapminder)
names(gapminder)
ncol(gapminder)
nrow(gapminder)
dim(gapminder)
```
## Do basic statistical and visual checks for each variable. There are six variables in the data frame: country, continent, year, lifeExp, pop, gdpPercap.

### 1. country
```{r}
class(gapminder$country)
nlevels(gapminder$country)
levels(gapminder$country)
summary(gapminder$country)
```

### 2. continent
```{r}
class(gapminder$continent)
nlevels(gapminder$continent)
levels(gapminder$continent)
summary(gapminder$continent)
barplot(table(gapminder$continent))
```

### 3. year
```{r}
class(gapminder$year)
summary(gapminder$year)
```

### 4. lifeExp
```{r}
class(gapminder$lifeExp)
summary(gapminder$lifeExp)
hist(gapminder$lifeExp)
```

### 5. pop
```{r}
class(gapminder$pop)
summary(gapminder$pop)
hist(gapminder$pop)
```

### 6. gdpPercap
```{r}
class(gapminder$gdpPercap)
summary(gapminder$gdpPercap)
hist(gapminder$gdpPercap)
```

## And lastly, some basic plots to visualize basic relationships in the data.

```{r}
plot(lifeExp ~ log(pop), gapminder)
plot(log(gdpPercap) ~ log(pop), gapminder)
plot(lifeExp ~ continent, gapminder)
plot(lifeExp ~ country, gapminder)

```


